Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-04-29 code-tests.tex
edits done through 1-zone collapse test.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Double mach reflection}
\label{sec.tests.doublemach}
\red{(Brian)}

The double mach reflection test is a classic two-dimensional test of
hydrodynamic algorithms, and was originally described in
\citet{1984JCoPh..54..115W} \citep[and more recently
in][]{2008ApJS..178..137S}.  In this problem (shown in
Figure~\ref{fig.doublemach}), a shock is injected at an angle to a
reflecting surface, and a jet appears along the reflecting surface.
The ideal solution is self-similar, and the appearance of this
solution is highly sensitive to numerical diffusion.  If numerical
noise is present, a Kelvin-Helmholz instability develops along this
jet and breaks the self-similarity.

In the test problem shown in Figure~\ref{fig.doublemach}, a 2D
simulation with $960 \times 240$ cells was formed with a domain of $x
= [0, 4]$ and $y = [0, 1]$.  We use an ideal gas equation of state of
$\gamma = 1.4$, a pre-shock density of 1.4, and a pre-shock specific
internal energy of $2.5/1.4$ (all in arbitrary units).  A Mach 10
shock is initialized with a shock normal that is 30$^\circ$ from the
x-axis and an initial position on the lower boundary of x$ = 1/6$.
The lower y boundary and right x boundary are reflecting; the left x
boundary is inflowing, and the upper y boundary has a time-dependent
boundary condition that allows the shock to propagate into the domain
as if it extends to infinity.  The simulation starts at t$ = 0$ and
runs until t$ = 0.205$ (arbitrary units), at which point the rightmost
extent of the shock should be at roughly x$ = 3$.  In this run, we use
the direct Eulerian implementation of the piecewise parabolic hydro
method, with diffusion, flattening, and shock steepening all turned
on.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/DoubleMachTest.eps}
\caption{Double mach test.}
\label{fig.doublemach}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.4\textwidth]{figures/sedov-ppm-slice.eps}
\includegraphics[width=0.4\textwidth]{figures/sedov-zeus-slice.eps}
\caption{Density slice from the Sedov Blast Test at $t =
0.07$. Left-hand image shows the results from PPM ({\tt HydroMethod =
0}), right-hand image shows the results from Zeus ({\tt Hydromethod =
2}). Notably, the Zeus shock front has progressed less far than forthe
PPM run. This is due to energy loss when conserving internal, not
total, energy.}
\label{fig.sedov1}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{figures/sedov-profiles.eps}
\caption{Radially averaged profiles for the Sedov Blast Test at $t =
0.07$. Clockwise from top-left shows density, velocity, internal
energy and pressure.}
\label{fig.sedov2}
\end{center}
\end{figure}

The Sedov Blast Test \citep{Sedov1959} models an intense explosion,
initiated by depositing thermal energy into a homogenous distribution
of gas. The result is a strong spherical shock wave centered on the
point of energy injection.  This problem is a popular test of
astrophysical numerical codes for three reasons: Firstly, it is
particularly appropriate to astrophysics since it represents the
situation when a supernovae explosion occurs. Secondly, it has an
exact analytical solution whereby the shock front's radial position is
given by:

@@

\noindent where $E_0$ is the initial energy injection, $\rho_0$ is the
background density and $\alpha = 1.0$ for cylindrical symmetry and an
ideal gas with $\gamma = 1.4$. For the full derivation see
\citet{Sedov1959}. This solution makes it possible to assess how well
the code performs. Thirdly, the spherical shock is a challenging
problem for numerical codes (both grid and particle) since the shock
wave increases in size as the simulation progresses and its
symmetrical nature highlights any directional preferences that grid
codes can succumb to. The test presented here is the two-dimensional
version that is included in the Enzo distribution. The
three-dimensional results from this test, both for Enzo and three
other leading astrophysics codes, can be founds in \citet{Tasker2008}.

In the initial state, the box contains a homogenous distribution of
gas at a density of 1 (note that, in the absense of gravity, this
problem is scale-free and so without units). Thermal energy is
deposited into a single cell at the center of the box with $E_0 =
10.0$. The problem is completed in two dimensions with reflecting
boundary conditions (the default for Enzo) and uses a boxsize of side
$1$. For this problem, we selected a top grid of $100 \times 100$ and
four levels of refinement, placed based on shock location and the
slope in density and total energy (this corresponds to cell flagging
methods 1 and 3 in the Enzo parameter file). The exception to this
scheme is in the initial conditions where grids are placed directly
around the injection point. The results were assessed at $t = 0.07$,
which corresponding to a time just before the shock reaches the box
edge (see Figure~\ref{fig.sedov1}).

Figure~\ref{fig.sedov2} shows the radial profiles for the simulation
run with the PPM hydro-solver (blue dashed line) and the Zeus
hydro-solver (red dash-dot-dot line) together with the analytical
solution (black solid line). Clockwise from the top-left are density,
velocity, internal energy and pressure. PPM matches the analytical
solution extremely well for all quantities. However, Zeus' shock front
lags behind the analytical position. This can also be seen in the
slices shown in Figure~\ref{fig.sedov1}. The cause of this discrepancy
is that Zeus shows a substantial energy lost during the first few
timesteps and produces a diamond-shaped, rather than spherical,
shockfront during this time. After this, the code correctly conserves
energy but this intial energy loss remains clearly visible in the
position of the shock at $t = 0.07$. This problem was addressed
directly by \citet{Clarke2010}, who attributed the source of the issue
to this version of Zeus solving the internal, rather than total,
energy equation. In situations with strong energy gradients, this
choice caused an energy loss and the artificial viscosity produces the
direction-dependent shockfront shape. In their paper,
\citet{Clarke2010} presents results from an alternative version of
Zeus that conserves total energy. This problem is less marked for
smaller energy gradients and it should be noted that Zeus' stability
and speed make it a highly competitive choice, despite the
disagreements in this test.

@@

The Zel`dovich pancake \citep{1970A&A.....5...84Z} is particularly
relevant to cosmological simulations, because it includes many
features that are critical to structure formation: hydrodynamics,
expansion, and self gravity.  This problem represents an ideal,
isolated caustic formation, and is thus a useful proxy for much more
complicated structures in full 3-dimensional cosmological simulations
(such as the collapse of gas onto a cosmological halo or filament).

The initial conditions are simple, and we follow the prescription of
\citet{Anninos94}.  Assuming a flat cosmology, the density
perturbation is given by

@@

\begin{equation}
v(x_l) = -H_0 \frac{1 + z_c}{(1+z)^{1/2}} \frac{\sin(k x_l}{k}
\end{equation}

In the equations above, $\rho_0$ is the background density, z$_c$ is a
free parameter and is the redshift at which the sheet forms a caustic
(i.e., 'pancakes'), z is the redshift of initialization, x$_l$ is the
Lagrangian mass coordinate, $k = 2 \pi / \lambda$ (where $\lambda$ is
the perturbation wavelength), and H$_0$ is the $z = 0$ value of the
Hubble constant.  Note that this solution is expressed in terms of
Lagrangian positions, so one needs to convert this into the Eulerian
coordinates, x$_e$, that are more useful to \enzo:

@@

We note that the solution described above is exact up to the point of
caustic formation.  In Figure~\ref{fig.pancake}, we show the results
of a test of the AMR version of \enzo's Zel'dovich Pancake test.  A
one-dimensional box of length $64$~Mpc/h is initialized at z$ = 20$ in
an $\Omega_M = 1$ universe with h$ = 0.5$ and a background temperature
of 100~K, $\rho_0 = \rho_c$.  The simulation is initialized with 64
grid cells, refining by factors of four using criteria based on cell
mass and the presence of shocks, for a maximum of 2 levels (i.e., an
equivalent maximum resolution of 1024 grid cells).  The simulation was
evolved to z$ = 0$ using the PPM hydro method, with the final output
of the calculation being shown in Figure~\ref{fig.pancake}.

Figure~\ref{fig.pancake} shows the key features of this test problem.
The strong shocks and large density gradients are well-resolved, with
density and velocity jumps being well-delineated.  The key features of
this test problem can be resolved with far fewer cells -- simulations
including a mere 8 cells resolve the key features, as shown in
sections 3.3.4-3.3.5 of~\citet{1996PhDT........80B} -- but we choose a
higher resolution here for illustrative purposes.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/AMRZeldovichPancake.eps}
\caption{Zel`Dovich Pancake test shown at z$ = 0$, initialized in a
$\Omega_m = 1$ universe at z$ = 20$ on a one-dimensional grid having
64 cells, and further refined by factors of four based on cell mass
and the presence of shocks for up to two additional levels of mesh,
having a maximum effective resolution of 1024 grid cells.}
\label{fig.pancake}
\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{MHD: Orszag-Tang Vortex}
\label{sec.tests.mhd}
Figure \ref{fig.orszag} shows the Orszag-Tang vortex problem, a
typical test problem for MHD \citep{Orszag79}.  The left panel shows
the MHD-CT result, while the right panel shows the Dedner result.
This test shows that small scale structure can be generated in MHD
from large scale initial perturbations.  The test begins with uniform
density, $\rho_0=25/36 \pi$ and pressure, $P_0=5/12 \pi$.  There is a
single rotational mode in the velocity, and two in the magnetic field:
${\bf v}_0 $ = (-sin(2$\pi$ y) $ \hat{x}$ , sin(2$\pi$ x) $\hat{y}$),
${\bf B}_0$ = (-sin(2 $\pi$ y) $ \hat{x},$ sin( 4 $\pi$ x )$\hat{y}$).
The simulation is evolved to t=0.48.  One can see that the structures
are accurately represented as compared to, for instance,
\citet{Toth00}, and that the resolution of shocks is comparable in
both methods.

{\emph Dave's note: I'll probably play around with these results in
future versions.}

\begin{figure}
\begin{center}
\includegraphics[width=0.4\textwidth]{figures/MHDCT_OrszagTang_Density.eps}
\includegraphics[width=0.4\textwidth]{figures/MHDDedner_OrszagTang_Density.eps}
\caption{TANG}
\label{fig.orszag}
\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{One-zone collapse test}
\label{sec.tests.1-zone}
The one-zone collapse test simulates the collapse of a
self-gravitating gas cloud using a semi-analytic model for the
evolution of gas density and adiabatic heat input as a function of
time.  It is designed to test the chemistry and cooling modules over a
wide range in densities and over physically motivated timescales.
Because this test disables the hydrodynamic and gravity solvers and
uses a simple model for the density evolution, it is far faster than
running a true collapse simulation.  The density evolution is based on
the self-similar Larson-Penston \citep{1969MNRAS.145..271L,
  1969MNRAS.144..425P} solution for isothermal collapse with a
modification to account for the efficiency with which the heat
introduced by compression can be radiated away
\citep{1983ApJ...265.1047Y}.  Our implementation, described briefly
here, follows the work of \citet{2005ApJ...626..627O}, which one
should consult for further details.  The density evolution is given by 
\begin{equation}
\frac{d\rho}{dt} = \frac{\rho}{t_{col}},
\end{equation}
where the collapse timescale, $t_{col}$, is
\begin{equation} \label{eqn.tcol}
t_{col} = \frac{t_{dyn}}{\sqrt{1 - f}},
\end{equation}
and $t_{dyn}$ is the dynamical time, expressed as
\begin{equation}
t_{dyn} = \sqrt{\frac{3 \pi}{32 G \rho}}.
\end{equation}
The collapse timescale is altered from the dynamical time by the
factor, $f$ in Equation \ref{eqn.tcol}, which an approximation of the
ratio of the pressure to the force of gravity.  The value of $f$
depends on the effective adiabatic index, $\gamma_{ef} \equiv (\partial
ln\ p / \partial ln\ \rho)$, which we calculate by linearly extrapolating
from derivative values for the two previous timesteps.  For the value
of $f$, we use the piecewise function of \citet{2005ApJ...626..627O},
given by
\begin{equation}
f = \left\{
  \begin{array}{ll}
  0, & \gamma_{ef} < 0.83,\\
  0.6 + 2.5 (\gamma_{ef} - 1) - 6.0 (\gamma_{ef} - 1)^{2}, & 0.83 <
  \gamma_{ef} < 1,\\
  1.0 + 0.2 (\gamma_{ef} - 4/3) - 2.9 (\gamma_{ef} - 4/3)^{2}, & \gamma_{ef} > 1.
\end{array} \right.
\end{equation}
The specific energy evolves as
\begin{equation}
\frac{de}{dt} = -p \frac{d}{dt} \frac{1}{\rho} - \Lambda,
\end{equation}
where $\Lambda$ is the cooling rate in units of erg s$^{-1}$ g$^{-1}$
and energy, temperature, density, and pressure are related by the
ideal gas law.  Figure \ref{fig.onezone} shows an example of the
one-zone collapse test performed with an initial number density of 1
cm$^{-3}$ and temperature of 100 K using the 12 species chemistry
network with H, D, and He species and metal cooling rates calculated
with Cloudy.

\begin{figure}
  \begin{center}
    \includegraphics[width=1.0\textwidth]{figures/OneZoneCollapseTest.eps}
  \end{center}
  \caption{Evolution of temperature versus number density for a one-zone
    collapse test for gas at metallicities from 0 to 10$^{-2} Z_{\odot}$
    using the primordial chemistry network with tabulated metal cooling
    rates calculated with Cloudy.}
  \label{fig.onezone}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Double mach reflection}
\label{sec.tests.doublemach}

The double mach reflection test is a classic two-dimensional test of
hydrodynamic algorithms, and was originally described in
\citet{1984JCoPh..54..115W} \citep[and more recently
in][]{2008ApJS..178..137S}.  In this problem, shown in
Figure~\ref{fig.doublemach}, a shock is injected at an angle to a
reflecting surface (the -y boundary), and a jet appears along the
reflecting surface.  The ideal solution is self-similar, and the
appearance of this solution is highly sensitive to numerical
diffusion.  If numerical noise is present, a Kelvin-Helmholz
instability develops along this jet and breaks the self-similarity.

In the test problem shown in Figure~\ref{fig.doublemach}, a 2D
simulation with $960 \times 240$ cells was formed with a domain of $x
= [0, 4]$ and $y = [0, 1]$.  We use an ideal gas equation of state of
$\gamma = 1.4$, a pre-shock density of 1.4, and a pre-shock specific
internal energy of $2.5/1.4$ (all in arbitrary units).  A Mach 10
shock is initialized with a shock normal that is 30$^\circ$ from the
x-axis and an initial position on the lower boundary of x$ = 1/6$.
The lower y boundary and right x boundary are reflecting; the left x
boundary is inflowing, and the upper y boundary has a time-dependent
boundary condition that allows the shock to propagate into the domain
as if it extends to infinity.  The simulation starts at t$ = 0$ and
runs until t$ = 0.205$ (arbitrary units), at which point the rightmost
extent of the shock should be at roughly x$ = 3$.  In this run, we use
the direct Eulerian implementation of the piecewise parabolic hydro
method, with the diffusion, flattening, and shock steepening all
turned on.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/DoubleMachTest.eps}
\caption{Density field at the final time in the Double Mach test.  A
Mach 10 shock is injected into the domain with a shock normal that is
30$^\circ$ from the x-axis with a time-dependent +y boundary condition
that mimics a shock of infinite length.  The solution is self-similar,
with the jet and whorls at $x \simeq 2.5-3.0$ in this figure being very sensitive
to numerical diffusion.  Our results compare favorably to the
higher-order images from \citet{1984JCoPh..54..115W}.}
\label{fig.doublemach}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.4\textwidth]{figures/sedov-ppm-slice.eps}
\includegraphics[width=0.4\textwidth]{figures/sedov-zeus-slice.eps}
\caption{Density slice from the Sedov Blast test at $t = 0.07$. Left:
results using the Piecewise Parabolic Method hydro scheme.  Right: results using the \zeus\
hydro method. Notably, the \zeus\ shock front has progressed less far
than in the PPM run. This is due to energy loss when conserving only
internal, and not total, energy.}
\label{fig.sedov1}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{figures/sedov-profiles.eps}
\caption{Radially averaged profiles for the Sedov Blast test at $t =
0.07$. Clockwise from top-left shows density, velocity, internal
energy and pressure.  The black solid line shows the analytic
solution.  The blue dashed line shows the simulation using the PPM
method, and the red dot-dashed line using \zeus.  The \zeus\ result
substantially lags the true result due to total energy not being
explicitly converged.}
\label{fig.sedov2}
\end{center}
\end{figure}

The Sedov Blast Test \citep{Sedov1959} models an intense explosion,
initiated by depositing thermal energy into a homogenous distribution
of gas. The result is a strong spherical shock wave centered on the
point of energy injection.  This problem is a popular test of
astrophysical numerical codes for three reasons: First, it is
particularly appropriate to astrophysics since it represents the
situation when a supernovae explosion occurs. Second, it has an
exact analytical solution whereby the shock front's radial position is
given by:

@@

\noindent where $E_0$ is the initial energy injection, $\rho_0$ is the
background density and $\alpha = 1.0$ for cylindrical symmetry and an
ideal gas with $\gamma = 1.4$. For the full derivation see
\citet{Sedov1959}. This solution makes it possible to assess how well
the code performs.  Third, the spherical shock is a challenging
problem for numerical codes (both grid and particle) since the shock
wave increases in size as the simulation progresses and its
symmetrical nature highlights any directional preferences that grid
codes can succumb to. The test presented here is the two-dimensional
version that is included in the \enzo\ distribution. The
three-dimensional results from this test, both for \enzo\ and three
other leading astrophysics codes, can be founds in \citet{Tasker2008}.

In the initial state, the box contains a homogenous distribution of
gas at a density of 1 (note that, in the absense of gravity, this
problem is scale-free and thus without units). Thermal energy is
deposited into a single cell at the center of the box with $E_0 =
10.0$. The problem is completed in two dimensions with reflecting
boundary conditions (the default for \enzo) and uses a box size of side
$1$. For this problem, we selected a top grid of $100 \times 100$
cells and a maximum of four levels of refinement, placed based on
shock location and the slope in density and total energy. The
exception to this scheme is in the initial conditions, where grids are
placed directly around the injection point. The results were assessed
at $t = 0.07$, which corresponding to a time just before the shock
reaches the box edge (see Figure~\ref{fig.sedov1}).

Figure~\ref{fig.sedov2} shows the radial profiles for the simulation
run with the PPM hydro-solver (blue dashed line) and the \zeus\
hydro-solver (red dash-dot-dot line) together with the analytical
solution (black solid line). Clockwise from the top-left are density,
velocity, internal energy and pressure. PPM matches the analytical
solution extremely well for all quantities. However, the shock front
in the \zeus\ simulation lags behind the analytical position. This can
also be seen in the slices shown in Figure~\ref{fig.sedov1}. The cause
of this discrepancy is that \zeus\ shows a substantial energy loss
during the first few timesteps and produces a diamond-shaped, rather
than spherical, shockfront during this time. After this, the code
correctly conserves energy but this intial energy loss remains clearly
visible in the position of the shock at $t = 0.07$. This problem was
addressed directly by \citet{Clarke2010}, who attributed the source of
the issue to this version of \zeus\ solving the internal, rather than
total, energy equation. In situations with strong energy gradients,
this choice caused an energy loss and the artificial viscosity
produces the direction-dependent shockfront shape. In their paper,
\citet{Clarke2010} presents results from an alternative version of
\zeus\ that conserves total energy. This problem is less marked for
smaller energy gradients and it should be noted that the \zeus\ hydro
algorithm's stability and speed make it a highly competitive choice,
despite the disagreements in this test.

@@

The Zel`dovich pancake \citep{1970A&A.....5...84Z} is particularly
relevant to cosmological simulations, because it includes many
features that are critical to structure formation: hydrodynamics,
expansion, and self gravity.  This problem represents the formation of an ideal,
isolated caustic, and is thus a useful proxy for much more
complicated structures in full 3-dimensional cosmological simulations
(such as the collapse of gas onto a cosmological halo or filament).

The initial conditions are simple, and we follow the prescription of
\citet{Anninos94}.  Assuming a geometrically flat cosmology, the density
perturbation is given by

@@

\begin{equation}
v(x_l) = -H_0 \frac{1 + z_c}{(1+z)^{1/2}} \frac{\sin(k x_l)}{k}
\end{equation}

In the equations above, $\rho_0$ is the background density, z$_c$ is a
free parameter and is the redshift where the sheet forms a caustic
(i.e., where it `pancakes'), z is the redshift of initialization, x$_l$ is the
Lagrangian mass coordinate, $k = 2 \pi / \lambda$ (where $\lambda$ is
the perturbation wavelength), and H$_0$ is the $z = 0$ value of the
Hubble constant.  Note that this solution is expressed in terms of
Lagrangian positions, so one needs to convert this into the Eulerian
coordinates, x$_e$, that are more useful to \enzo:

@@

We note that the solution described above is exact up to the point of
caustic formation.  In Figure~\ref{fig.pancake}, we show the results
of a test of the adaptive mesh version of \enzo's Zel'dovich Pancake test.  A
one-dimensional box of length $64$~Mpc/h is initialized at z$ = 20$ in
an $\Omega_M = 1$ universe with h$ = 0.5$ and a background temperature
of 100~K, $\rho_0 = \rho_c$.  The simulation is initialized with 64
grid cells, refining by factors of four using criteria based on cell
mass and the presence of shocks, for a maximum of 2 levels (i.e., an
equivalent maximum resolution of 1024 grid cells).  The simulation was
evolved to z$ = 0$ using the PPM hydro method, with the final output
of the calculation being shown in Figure~\ref{fig.pancake}.

Figure~\ref{fig.pancake} shows the key features of this test problem.
The strong shocks and large density gradients are well-resolved, with
density and velocity jumps being well-delineated and at the correct locations.  The key features of
this test problem can be resolved with far fewer cells -- simulations
including a mere 8 cells resolve the key features, as shown in
Sections 3.3.4-3.3.5 of~\citet{1996PhDT........80B} -- but we choose a
higher resolution here for illustrative purposes.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/AMRZeldovichPancake.eps}
\caption{Zel`Dovich Pancake test shown at z$ = 0$, initialized in a
$\Omega_m = 1$ universe at z$ = 20$ on a one-dimensional grid having
64 cells, and further refined by factors of four based on cell mass
and the presence of shocks for up to two additional levels of mesh,
having a maximum effective resolution of 1024 grid cells. The top,
middle, and bottom rows show density, temperature, and velocity of the
gas, respectively; all are a function of position in units of the box
size.  The central region ($x \simeq 0.45-0.55$) has bene adaptively
refined, as can be seen by the locations of grid points.  Shocks and
the central density peak are clearly resolved, with well-delineated
jumps at the appropriate locations.}
\label{fig.pancake}
\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{MHD: Orszag-Tang Vortex}
\label{sec.tests.mhd}
Figure~\ref{fig.orszag} shows the Orszag-Tang vortex problem, a
standard magnetohydrodynamic test problem \citep{Orszag79}.  The left panel shows
the result using \enzo's constrained transport MHD method, while the right panel shows the result using
\enzo's implementation of Dedner MHD.
The Orszag-Tang vortex test shows that significant small scale structure can be generated in MHD
from large scale initial perturbations, and is used to compare the
effective resolution of different MHD schemes.  The test begins with uniform
density, $\rho_0=25/36 \pi$ and pressure, $P_0=5/12 \pi$ (as with
other tests in this section, in the absence of gravity or chemistry we use
arbitrary units).  There is a
single rotational mode in the velocity, and two in the magnetic field:
${\bf v}_0 $ = (-sin(2$\pi$ y) $ \hat{x}$ , sin(2$\pi$ x) $\hat{y}$),
${\bf B}_0$ = (-sin(2 $\pi$ y) $ \hat{x},$ sin( 4 $\pi$ x )$\hat{y}$).
The simulation is evolved to $t=0.48$ (in arbitrary units).  One can see that the structures
are accurately represented as compared to, for instance,
\citet{Toth00}, and that the resolution of shocks is comparable in
both methods.

\begin{figure}
\begin{center}
\includegraphics[width=0.4\textwidth]{figures/MHDCT_OrszagTang_Density.eps}
\includegraphics[width=0.4\textwidth]{figures/MHDDedner_OrszagTang_Density.eps}
\caption{Density field from the Orszag-Tang vortex test, at $t=0.48$.
Left: solution using constrained transport MHD.  Right: solution using
Dedner MHD. The initial conditions are uniform density, with a single
rotating velocity structure and two circular magnetic structures.
These initial conditions generate significant small-scale structure in
both the CT and Dedner schemes, which have approximately equal
effective resolution.}
\label{fig.orszag}
\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{One-zone collapse test}
\label{sec.tests.1-zone}
The one-zone collapse test simulates the collapse of a
self-gravitating gas cloud using a semi-analytic model for the
evolution of gas density and adiabatic heat input as a function of
time.  It is designed to test the chemistry and cooling modules over a
wide range in densities and over physically motivated timescales.
Because this test disables the hydrodynamic and gravity solvers and
uses a simple model for the density evolution, it is far faster than
running a true collapse simulation.  The density evolution is based on
the self-similar Larson-Penston \citep{1969MNRAS.145..271L,
  1969MNRAS.144..425P} solution for isothermal collapse with a
modification to account for the efficiency with which the heat
introduced by compression can be radiated away
\citep{1983ApJ...265.1047Y}.  Our implementation, described briefly
here, follows the work of \citet{2005ApJ...626..627O}, and we direct
the interested reader to this paper for further details.  The
evolution of the gas density, $\rho$, is given by 
\begin{equation}
\frac{d\rho}{dt} = \frac{\rho}{t_{col}},
\end{equation}
where the collapse timescale, $t_{col}$, is
\begin{equation} \label{eqn.tcol}
t_{col} = \frac{t_{dyn}}{\sqrt{1 - f}},
\end{equation}
In this equation, $t_{dyn}$ is the dynamical time and is expressed as
\begin{equation}
t_{dyn} = \sqrt{\frac{3 \pi}{32 G \rho}}.
\end{equation}

with G being the gravitational constant.  The collapse timescale is altered from the dynamical time by a
factor $f$ in Equation \ref{eqn.tcol}, which is an approximation of the
ratio of the gas pressure to the force of gravity.  The value of $f$
depends on the effective adiabatic index, $\gamma_{ef} \equiv (\partial
ln\ p / \partial ln\ \rho)$, which we calculate by linearly extrapolating
from derivative values for the two previous timesteps.  For the value
of $f$ in this test problem, we use the piecewise function of \citet{2005ApJ...626..627O},
given by
\begin{equation}
f = \left\{
  \begin{array}{ll}
  0, & \gamma_{ef} < 0.83,\\
  0.6 + 2.5 (\gamma_{ef} - 1) - 6.0 (\gamma_{ef} - 1)^{2}, & 0.83 <
  \gamma_{ef} < 1,\\
  1.0 + 0.2 (\gamma_{ef} - 4/3) - 2.9 (\gamma_{ef} - 4/3)^{2}, & \gamma_{ef} > 1.
\end{array} \right.
\end{equation}
The specific energy evolves as
\begin{equation}
\frac{de}{dt} = -p \frac{d}{dt} \frac{1}{\rho} - \Lambda,
\end{equation}
where $\Lambda$ is the cooling rate in units of erg s$^{-1}$ g$^{-1}$
and energy, temperature, density, and pressure are related by the
ideal gas law.  Figure \ref{fig.onezone} shows an example of the
one-zone collapse test performed with an initial number density of 1
hydrogen atom per
cm$^{-3}$ and an initial temperature of 100 K using the 12 species chemistry
network with H, D, and He species and metal cooling rates calculated
with the \texttt{Cloudy} code.  The effects of metal cooling can be
clearly seen; as the metallicity increases from zero to
$10^{-2}$~Z$_\odot$, the gas deviates from the primoridal result
(black line) both more rapidly and more significantly.  Our primordial results
compare very well to those shown in \citet{2005ApJ...626..627O}; however,
we use a different cooling method than Omukai, so the lines describing
the evolution of the metal-enriched gas are not directly comparable.

\begin{figure}
  \begin{center}
    \includegraphics[width=1.0\textwidth]{figures/OneZoneCollapseTest.eps}
  \end{center}
  \caption{Evolution of temperature versus number density for a one-zone
    collapse test for gas at metallicities from 0 to 10$^{-2}
    Z_{\odot}$.  This test is based on the results of
    \citet{2005ApJ...626..627O}, and approximates the collapse of a
    self-gravitating, cooling gas cloud.  This test problem uses \enzo's primordial chemistry network with tabulated metal cooling
    rates calculated with the \texttt{Cloudy} code, and compares
    favorably to Omukai's results.}
  \label{fig.onezone}
\end{figure}
